Accessing the nucleon tensor structure in inclusive deep inelastic scattering

### 2016-07-26 1607 - JMC for g2/Paper/JMC_dis.tex
Some corrections

The correlator $\Xi$ can be parametrized in terms of jet parton correlation
functions (PCFs), using the vectors $l$ and $n_+$: 
\begin{equation}
\Xi(l,n_+) = \Lambda A_1(l^2)\,{\bm 1} + A_2(l^2)\,\lslash 
+ \frac{\Lambda^2}{l \cdot n_{+}} \nslash_+ \, B_{1}(l^2)
+ \frac{i \Lambda}{2 l \cdot n_{+}} [\,\lslash,\nslash_+ ] \, B_{2}(l^2) \ .
\label{e:jetexpansion}
\end{equation} 
Time reversal invariance in QCD requires $B_{2}=0$, while $B_{1}$ contributes
only at twist-4 order, and will not be considered further in this paper. We
focus, instead, on the role of chiral odd terms in the $g_2$ structure function up to twist 3. At this order, 
\begin{equation}
  \Xi(l,n_+) = \Lambda A_1(l^2)\,{\bm 1} + A_2(l^2)\,\lslash 
    + O(\Lambda^2/Q^2)
\label{e:jetexpansion-tw3}
\end{equation} 
is nothing else than the full quark propagator; note however, that we consider
here the full QCD vacuum rather than the perturbative one.  
The $A_1$ and $A_2$ terms can be nicely interpreted in terms of the spectral representation of the cut quark propagator \cite{D'Hoker-lecture-notes,Romao-lecture-notes},
\begin{align} 
  \Xi(l) =  
  \int d \mu^2 \big[ J_1 (\mu^2)\,\mu + J_2 (\mu^2)\,\lslash \big] \,
  \delta(l^2 -\mu_j^2) \ ,
\label{e:jetspectral}
\end{align}
where $\mu^2$ is interpreted as the invariant mass of the current jet, {\it
  i.e.}, of the particles going through the cut in the top blob of
Fig.\ref{fig:handbags}(a), and the $J_i$ are the spectral functions of the
quark propagator, that have been also called ``jet functions'' in
\cite{Accardi:2008ne,SCET?}. These satisfy
\begin{align}
  J_2(\mu^2) \geq J_1(\mu^2) \geq 0
  \hspace*{0.5cm} \text{and} \hspace*{0.5cm}
  \int d\mu^2 J_2(\mu^2) = 1 \ .
\label{eq:jetfnsprops}
\end{align}
From a comparison of Eqns.\eqref{e:jetexpansion} and \eqref{e:jetspectral}, one can see that 
\begin{align}
  A_1(l^2)&=\frac{\sqrt{l^2}}{\Lambda}J_1(l^2) & A_2(l^2)&=J_2(l^2) \ .
  \label{eq:jet_vs_spectral}
\end{align}

@@

\cite{Weinberg:1995mt}, while $0 < \mj < \int d\mu^2 \mu J_2(\mu^2)$ is

@@

  \mj = O(10-100 \text{ MeV}) \ .

@@

Calculating this on the perturbative vacuum and limiting oneself to LO

@@

The novel element in our analysis is the jet's quark-gluon-quark correlator
$\Xi_A^{\mu}(l,k)$ in diagrams \ref{fig:handbags}(c), 
\begin{equation} 
\begin{split} 
  \left(\Xi_A^{\mu} \right)_{ij} &=
   \frac{1}{2}\, \sum_X \int \frac{\de \eta^+\, \de^2 \bm{\eta}_T}{(2\pi)^{3}}\;
   e^{\ii k \cdot \eta}\,
   \langle 0|\,
   {\cal U}^{n_+}_{(+\infty,\eta)}\,
   g A^{\mu}(\eta)\,
   \,\psi_i(\eta)|X\rangle
   \langle X|
             \bar{\psi}_j(0)\,
   {\cal U}^{n_+}_{(0,+\infty)}
   |0\rangle \bigg|_{\eta^-=0} .
\label{e:xi_A}
\end{split} 
\end{equation}  
This diagram and its hermitian conugate are not only important to account for
all contribution of order $O(1/Q)$, but also in restoring up to twist-3 the
gauge invariance broken in diagram \ref{fig:handbags}(a) by the different mass
of the incoming and outgoing quark lines, namely, $\mq \neq \mj$.

@@

% As we shall see, an analogous formula for quark-gluon-quark correlators is not needed.
%The $\xi_1$ term is chiral-odd and therefore can appear in the inclusive cross section only coupled to the transversity function $h_1$. Therefore, for our analysis 
The relevant part of the semi-inclusive hadronic tensor for our analysis is 
\begin{align}
  \label{eq:Wsidis_ini}
  2 \Lambda  W^{\mu\nu}
    & = i \frac{2\Lambda}{Q} \hat t^{[\mu}_{\phantom \perp} 
    \epsilon_\perp^{\nu]\rho}S_{\perp\rho} 
    \sum_q e_q^2
    \bigg[ 2 x_b g_T(x_B) \sum_h \int dz d^2p_{hT} D_1^{q,h}(z,p_{hT}) 
  %\\ &
  + 2 h_1(x_B) \sum_h \int dz d^2p_{hT} \tilde E^{q,h}(z,p_{hT}) \bigg] + \ldots
\end{align}
Here and we reintroduced the quark flavor $q$ for maximum clarity, and $e_q$ its electric charge.
The first term can be easily integrated with the help of the sum rule
\eqref{eq:SIDIS_to_DIS_TMDlevel}. To integrate the latter, we first need make
use of the relation $\tilde E(z) = E(z) - (\mq/\Lambda) z D_1(z)$, which is a
consequence of the QCD equations of motion \cite{Bacchetta:2006tn}, then
utilize the sum rule \eqref{eq:SIDIS_to_DIS}: 
\begin{align}
  \sum_h \int dz d^2p_{hT} \tilde E^{q,h}(z,p_{hT}) 
    = \sum_h \int dz d^2p_{hT} \Big[ E^{q,h}(z,p_{hT}) - \frac{\mq}{\Lambda} z D_1^{q,h}(z,p_{hT}) \Big]
    = \xi_1 - \frac{\mq}{\Lambda} \xi_2 = \frac{\mj - \mq}{\Lambda} \ .
\end{align}
This formula is the single most important result of this paper, and provides a
non perturbative generalization of the commonly used $\int\tilde E =0$ sum
rule introduced in \cite{Jaffe:1996zw}. Indeed, calculating the jet correlator 
on the perturbative vacuum one would obtain, as already discussed, $\mj=\mq$
and the new term would vanish.

Finally, with suitable projections of the hadronic tensor, the inclusive cross section up to order $\Lambda/Q$ can be written as
\begin{align}
\frac{d\sigma}{d\xbj \, dy\, d\psi}
%&
=
\frac{2 \alpha^2}{\xbj y Q^2}\,
\frac{y^2}{2\,(1-\varepsilon)}\, 
\biggl\{
&F_{UU ,T} + \varepsilon F_{UU ,L}
+ S_\parallel \lambda_e\,
  \sqrt{1-\varepsilon^2}\; 
F_{LL}
%\nonumber 
%\\  &
+ |\bm{S}_\perp| \lambda_e\, \sqrt{2\,\varepsilon (1-\varepsilon)}\, 
  \cos\phi_S\, 
F_{LT}^{\cos \phi_S}
 \biggr\} \ ,
\label{e:crossdis}
\end{align}
where the structure functions on the right hand side are defined as
\begin{align}
F_{UU ,T} &= \xbj\,\sum_q e_q^2\,f_1^q(\xbj),
\\
F_{UU ,L} &= 0,
\\
F_{LL} &=\xbj\,\sum_q e_q^2\,g_1^q(\xbj),
\\
F_{UT}^{\sin \phi_S}&=0,
\label{e:FUTint}
\\
F_{LT}^{\cos \phi_S}&=-\xbj\,\sum_q e_q^2\, \frac{2\Lambda}{Q}\,
\biggl(\xbj  g_T^q(\xbj)
   + \frac{\mj -\mq}{\Lambda} \, h_{1}^q(\xbj) \biggr).
\label{e:FLTint}
\end{align}
The second term in the last structure function is a new result from our
analysis; it is not suppressed as an inverse power of $Q$, and therefore
survives even in the Bjorken limit. On the non-perturbative
vacuum the jet mass is larger than the quark's, and this contributes a
non-negligible term to the twist-3 part of the $g_2$ function, as we will
discuss in the next section.

@@

The new term in Eq.\eqref{e:FLTint} only appears in the $g_2$ structure function. Following the derivation in Ref.~\cite{Accardi:2009au}, one finds
\begin{align}
\label{e:g2}
  g_2(x_B) = g_2^{WW} + \frac{1}{2}\,\sum_a e_a^2
\biggl(
    \widetilde g_T^{a \star}(x) 
    + \int_x^1\frac{dy}{y} \widehat{g}_T^q(y) 
    + \frac{\mq}{\Lambda} \left(\frac{h_1^q}{x}\right)^\star(x) 
    + \frac{\mj-\mq}{\Lambda} \frac{h_1^q(x)}{x} 
\Biggr) \ ,
\end{align}
where we defined $f^*(x) = -f(x) + \int_x^1\frac{dy}{y} f(y)$. The first 4
terms coincide with the result obtained in the conventional handbag
approximation \cite{Accardi:2009au}, while the fifth is new.

The first term is also known as the Wandzura-Wilczeck function $g_2^{WW} =
-g_1^*(x)$ , and contains all the ``pure twist-2'' chiral even contributions to
the $g_2$ structure coming from quark-quark correlators. The second and third
terms contain all ``pure twist-3'' contributions, i.e., those coming from
quark-gluon-quark correlators. The fourth and fifth terms depend on the
transversity parton distribution function, $h_1$. 
The fourth term is usually neglected for
light quarks since it is proportional to $\mq=O$(1 MeV). In the last term,
new in our analysis, the transversity distribution is multiplied by a constant
of $O$(100 MeV), and cannot be a priori neglected.

It is important to estimate the size of the various contributions to the non Wandzura-Wilczek part of $g_2$. We define the shorthand notation
\begin{align}
g_2^{tw3} & = \frac{1}{2}\,\sum_a e_a^2
    \biggl(
    \widetilde g_T^{a \star}(x) 
    + \int_x^1\frac{dy}{y} \widehat{g}_T^q(y) 
    \biggr) 
&
g_2^{\text{quark}} &= \frac{1}{2}\,\sum_a e_a^2 
 \frac{\mq}{\Lambda} (h_1^q/x)^\star(x),
&
g_2^{\text{jet}} &= \frac{1}{2}\,\sum_a e_a^2 
\frac{\mj-\mq}{\Lambda} \frac{h_1^q(x)}{x}. 
\end{align} 
These are comapred in Figure~\ref{f:g2contrib} to the $g_2-g_2^{WW}$ function
obtained in the very recent JAM15 fit of polarized DIS asymmetries
\cite{Sato:2016tuz}, that includes a large amount of precise data at large $x$
from Jefferson Lab, and simultaneously fits the higher-twist components of in
$g_1$ and $g_2$ to the data. For the ``pure twist-3'' contribution,
$g_2^{tw3}$, {\it i.e.}, the contribution from quark-gluon-quark matrix
elements, we show a model calculation by Braun et al.~\cite{Braun:2011aw}; for
(modified) bag model calculations, see \cite{Jaffe:1990qh,Stratmann:1993aw}. To estimate the contributions
from quark ($g_2^q$) and jet mass ($g_2^{jet}$) effects, that depend on chiral
odd quark-quark matrix elements, we use the recent Pavia15 fit of the
transversity distribution from Ref.~\cite{Radici:2015mwa}, which is comparable
also to other 
extractions~\cite{Anselmino:2013vqa,Kang:2015msa}. Furthermore, we choose the
values of the mass parameters to be $\mq=5$ MeV and $\mj = 100$ MeV.

As one can see, in the proton case the pure twist-3 contribution is quite
smaller in magnitude, and opposite in sign, compared to the JAM15 fit. As
expected, the quark-mass contribution is essentially negligible and cannot
reconcile these two. Even though the uncertainties in the $h_1$ extraction, and
even more in the estimate of $M_q$, are large, it is quite clear that the gap
between the pure twist-3 $g_2^{tw3}$ function and the JAM15 fit can be
explained by the new jet-mass contribution we discussed in this paper.

In the neutron case, the jet contribution is very negative at intermediate to
large values of $x$. If one trusts the order of magnituse of the $g_2^{tw3}$
calculation by Braun et al., one would conclude that the jet contribution
cannot be that large. However, the latter is dominated by the $d$ quark's
transversity, whose fit suffers from large systematic uncertainties and
saturates the negative Soffer bound.Recent deata in $p+p$ collisions indicate,
however, that $h_1^d$ is less negative than in the Pavia15 fits, in agreement
with the JAM15 fit of the non Wandzura-Wilczek contribution to
$g_2$. Correspondingly the jet contribution to the proton at $x \approx 0.1$
would become less positive, inproving as well the agreement with the JAM15
fit.

\begin{itemize}
\item smaller-$x$: constrains the small-$x$ behavior of transversity.
\end{itemize}
Some ideas to check:
\begin{itemize}
\item JAM13 neutron has almost 0 twist 3, with small positive contribution at
  small x. Can we use this to highlight the $g_2^{jet}$ contribution? Or maybe
  as a cnstrain: if Braun is right then $g_2^{jet}(n) \approx
  -g_2^{tw3}(n)$. (still, does not help much for the proton). 
\item In Jam 13, Fig. 6, it looks like in general the violation of the BC sum
  rule is a small-x effect --> constraints on h(x) fits? This cannot really be
  checked vs. JAM15, that assumes $g_2^(0)$ respects BC.  
\end{itemize}

@@

The zero-th moment provides an interesting relationship between the longitudinal and tensor spin structure of the target,
\begin{align}
  \label{eq:BC}
  \int dx\, g_2(x) = \sum_q e_q^2 \frac{\mj-\mq}{\Lambda} \int dx\, \frac{1}{x} h_1^q(x) \ ,
\end{align}
where we used the fact that $\hat g_T^q[0]=0$ identically due to the symmetry properties of the quark-gluon-quark correlators. The sum rule \eqref{eq:BC}, in which the pure twist-3 part does not take part, is a consequence of the Lorentz invariance of QCD interactions, that entails $\int_0^1 dx g_1^a(x) = \int_0^1 g_T^q(x)$. It explicitly displays a contribution from spin-flip processes, that are included in the original derivation by Burkhardt and Cottingham \cite{Burkhardt:1970ti} but do not show up in treatments that only consider free field quark propagators for the struck quark \cite{Jaffe:1996zw}. Since $h_1$ is driven to 0 by QCD evolution as $Q^2 \to \infty$, the BC sum rule $\int_0^1 dx\, g_2(x) =0$ is satisfied at least asymptotically. At finite scales, assuming the Burkhardt-Cottingham sum rule, that is derived at the cross section level (under suitable assumptions on the high-energy behavior of te hspin-flip amplitude) rather than from an analysis of collinear correlators, as done here, to be strictly valid, we obtain a strong constraint on the transversity function,
\begin{align}
   \int dx\, \frac{1}{x} h_1^q(x) = 0 \ .
\label{eq:ABsumrule}
\end{align}
Interestingly, one can show that this constraint, if valid at any given scale $Q_0$ is conserved through QCD evolution. Even if the BC sum rule is broken either by a $J = 0$ fixed pole with non-polynomial residue \cite{Jaffe:1996zw} or explicitly as, \eg, in the quark target model \cite{Metz-Kundu} -- i.e., if $g_2$ integrates to a finite but non-zero number -- we obtain that $h_1(x)/x$ must be integrable. This entails a bound on the small $x$ behavior of the transversity,
\begin{align}
  h_1^q(x) \propto x^\epsilon \ \ \ \epsilon>0 \ .
\label{eq:ABbound}
\end{align}
This bound will be very useful, \eg, in transversity fits, where the data at
small $x$ is, as yet, very limited, and in general for proper extrapolations
when calculating moments.

@@

\begin{figure}[tbh]
  \centering
  \parbox[b]{8cm}{\centering
  \includegraphics
      [width=7cm,clip=true]
      {epl_emn_h_X.eps}
  }
  \parbox[b]{8cm}{\vskip-0.5cm\centering
  \includegraphics
      [width=7cm,clip=true]
      {epl_emn_h_h_X.eps}
  }
  \caption{\small
	Single hadron {\it (left)} and double hadron {\it (right)}
	production in $e^+ e^-$ collisions at LO with jet and
	fragmentation correlators.
  }
  \label{fig:epl_emn_h}
\end{figure}

>>>>

The correlator $\Xi$ can be parametrized in terms of jet parton correlation
functions (PCFs), using the vectors $l$ and $n_+$: 
\begin{equation}
\Xi(l,n_+) = \Lambda A_1(l^2)\,{\bm 1} + A_2(l^2)\,\lslash 
+ \frac{\Lambda^2}{l \cdot n_{+}} \nslash_+ \, B_{1}(l^2)
+ \frac{i \Lambda}{2 l \cdot n_{+}} [\,\lslash,\nslash_+ ] \, B_{2}(l^2) \ .
\label{e:jetexpansion}
\end{equation} 
Time reversal invariance in QCD requires $B_{2}=0$, while $B_{1}$ contributes
only at twist-4 order, and will not be considered further in this paper. We
focus, instead, on the role of chiral odd terms in the $g_2$ structure function up to twist 3. At this order, 
\begin{equation}
  \Xi(l,n_+) = \Lambda A_1(l^2)\,{\bm 1} + A_2(l^2)\,\lslash 
    + O(\Lambda^2/Q^2)
\label{e:jetexpansion-tw3}
\end{equation} 
is nothing else than the full quark propagator; note however, that we consider
here the full QCD vacuum rather than the perturbative one.  
The $A_1$ and $A_2$ terms can be nicely interpreted in terms of the spectral
representation of the cut quark propagator (see, e.g., Sec.~6.3 of
\cite{D'Hoker:2004aa} and Sec.~2.7.2 of \cite{Romao:2013aa}),
\begin{align} 
  \Xi(l) =  
  \int d \mu^2 \big[ J_1 (\mu^2)\,\mu + J_2 (\mu^2)\,\lslash \big] \,
  \delta(l^2 -\mu_j^2) \ ,
\label{e:jetspectral}
\end{align}
where $\mu^2$ is interpreted as the invariant mass of the current jet, {\it
  i.e.}, of the particles going through the cut in the top blob of
Fig.\ref{fig:handbags}(a), and the $J_i$ are the spectral functions of the
quark propagator, that have been also called ``jet functions'' in
\cite{Accardi:2008ne}. These satisfy
\begin{align}
  J_2(\mu^2) \geq J_1(\mu^2) \geq 0
  \hspace*{0.5cm} \text{and} \hspace*{0.5cm}
  \int d\mu^2 J_2(\mu^2) = 1 \ .
\label{eq:jetfnsprops}
\end{align}
From a comparison of Eqns.\eqref{e:jetexpansion} and \eqref{e:jetspectral}, one can see that 
\begin{align}
  A_1(l^2)&=\frac{\sqrt{l^2}}{\Lambda}J_1(l^2) & A_2(l^2)&=J_2(l^2) \ .
  \label{eq:jet_vs_spectral}
\end{align}

@@

(see Sec.~10.7 of Ref.~\cite{Weinberg:1995mt}), while $0 < \mj < \int d\mu^2 \mu J_2(\mu^2)$ is

@@

  \mj = O(10^2 \text{ MeV}) \ .

@@

Calculating this on the perturbative vacuum and to leading order

@@

The novel element in our analysis is the jet's quark-gluon-quark correlator
$\Xi_A^{\mu}(l,k)$ in diagrams \ref{fig:handbags}(c), 
\begin{equation} 
\begin{split} 
  \left(\Xi_A^{\mu} \right)_{ij} &=
   \frac{1}{2}\, \sum_X \int \frac{\de \eta^+\, \de^2 \bm{\eta}_T}{(2\pi)^{3}}\;
   e^{\ii k \cdot \eta}\,
   \langle 0|\,
   {\cal U}^{n_+}_{(+\infty,\eta)}\,
   g A^{\mu}(\eta)\,
   \,\psi_i(\eta)|X\rangle
   \langle X|
             \bar{\psi}_j(0)\,
   {\cal U}^{n_+}_{(0,+\infty)}
   |0\rangle \bigg|_{\eta^-=0} .
\label{e:xi_A}
\end{split} 
\end{equation}  
This diagram and its Hermitian conjugate are not only important to account for
all contribution of order $O(1/Q)$, but also to restore the
gauge invariance, which is broken in diagram \ref{fig:handbags}(a) due to the different mass
of the incoming and outgoing quark lines, namely, $\mq \neq \mj$.

@@

% As we shall see, an analogous formula for quark-gluon-quark correlators is not needed.
%The $\xi_1$ term is chiral-odd and therefore can appear in the inclusive cross section only coupled to the transversity function $h_1$. Therefore, for our analysis 
The relevant part of the semi-inclusive hadronic tensor for our analysis is 
\begin{align}
  \label{eq:Wsidis_ini}
  2 \Lambda  W^{\mu\nu}
    & = i \frac{2\Lambda}{Q} \hat t^{[\mu}_{\phantom \perp} 
    \epsilon_\perp^{\nu]\rho}S_{\perp\rho} 
    \sum_q e_q^2
    \bigg[ 2 x_b g_T(x_B) \sum_h \int dz d^2p_{hT} D_1^{q,h}(z,p_{hT}) 
  %\\ &
  + 2 h_1(x_B) \sum_h \int dz d^2p_{hT} \tilde E^{q,h}(z,p_{hT}) \bigg] + \ldots
\end{align}
For clarity, here we reintroduced the quark flavor $q$, $e_q$ being its electric charge.
The first term can be easily integrated with the help of the sum rule
\eqref{eq:SIDIS_to_DIS_TMDlevel}. To integrate the latter, we first need make
use of the relation $\tilde E(z) = E(z) - (\mq/\Lambda) z D_1(z)$, which is a
consequence of the QCD equations of motion \cite{Bacchetta:2006tn}, then
make use of the sum rule \eqref{eq:SIDIS_to_DIS}: 
\begin{align}
  \sum_h \int dz d^2p_{hT} \tilde E^{q,h}(z,p_{hT}) 
    = \sum_h \int dz d^2p_{hT} \Big[ E^{q,h}(z,p_{hT}) - \frac{\mq}{\Lambda} z D_1^{q,h}(z,p_{hT}) \Big]
    = \xi_1 - \frac{\mq}{\Lambda} \xi_2 = \frac{\mj - \mq}{\Lambda} \ .
\end{align}
This formula is the single most important result of this paper, and provides a
non perturbative generalization of the commonly used $\int\tilde E =0$ sum
rule introduced in \cite{Jaffe:1996zw}. Indeed, calculating the jet correlator 
on the perturbative vacuum one would obtain, as already discussed, $\mj=\mq$
and the new term would vanish.

Finally, the contraction of the hadronic tensor with the leptonic tensor leads
to the following well known result for the inclusive DIS cross section up to order $\Lambda/Q$~\cite{Bacchetta:2006tn}
\begin{align}
\frac{d\sigma}{d\xbj \, dy\, d\psi}
%&
=
\frac{2 \alpha^2}{\xbj y Q^2}\,
\frac{y^2}{2\,(1-\varepsilon)}\, 
\biggl\{
&F_{UU ,T} + \varepsilon F_{UU ,L}
+ S_\parallel \lambda_e\,
  \sqrt{1-\varepsilon^2}\; 
F_{LL}
%\nonumber 
%\\  &
+ |\bm{S}_\perp| \lambda_e\, \sqrt{2\,\varepsilon (1-\varepsilon)}\, 
  \cos\phi_S\, 
F_{LT}^{\cos \phi_S}
 \biggr\} \ ,
\label{e:crossdis}
\end{align}
where the structure functions on the right hand side correspond to
\begin{align}
F_{UU ,T} &= \xbj\,\sum_q e_q^2\,f_1^q(\xbj),
\\
F_{UU ,L} &= 0,
\\
F_{LL} &=\xbj\,\sum_q e_q^2\,g_1^q(\xbj),
\\
F_{UT}^{\sin \phi_S}&=0,
\label{e:FUTint}
\\
F_{LT}^{\cos \phi_S}&=-\xbj\,\sum_q e_q^2\, \frac{2\Lambda}{Q}\,
\biggl(\xbj  g_T^q(\xbj)
   + \frac{\mj -\mq}{\Lambda} \, h_{1}^q(\xbj) \biggr).
\label{e:FLTint}
\end{align}
The second term in the last structure function is a new result from our
analysis; it is not suppressed as an inverse power of $Q$ compared to the
standard term. On the nonperturbative
vacuum the jet mass is larger than the quark's, and this contributes a
nonnegligible term to the twist-3 part of the $g_2$ function, as we will
discuss in the next section.

@@

The new term in Eq.\eqref{e:FLTint} only appears in the $g_2$ structure function. Following the derivation in Ref.~\cite{Accardi:2009au}, one finds
\begin{align}
\label{e:g2}
  g_2(x_B) = g_2^{WW} + \frac{1}{2}\,\sum_a e_a^2
\biggl(
    \widetilde g_T^{a \star}(x) 
    + \int_x^1\frac{dy}{y} \widehat{g}_T^q(y) 
    + \frac{\mq}{\Lambda} \left(\frac{h_1^q}{x}\right)^\star(x) 
    + \frac{\mj-\mq}{\Lambda} \frac{h_1^q(x)}{x} 
\Biggr) \ ,
\end{align}
where we defined $f^*(x) = -f(x) + \int_x^1\frac{dy}{y} f(y)$. The first four
terms coincide with the result obtained in the conventional handbag
approximation \cite{Accardi:2009au}, while the fifth is new.

The first term is also known as the Wandzura-Wilczeck function $g_2^{WW} =
-g_1^*(x)$ , and contains all the twist-2 chiral-even contributions to
the $g_2$ structure coming from quark-quark correlators. The second and third
terms contain all ``pure twist-3'' contributions, i.e., those coming from
quark-gluon-quark correlators. The fourth and fifth terms depend on the
transversity parton distribution function, $h_1$. 
The fourth term is usually neglected for
light quarks since it is proportional to $\mq=O$(1 MeV). In the last term,
new in our analysis, the transversity distribution is multiplied by a constant
of $O$(100 MeV), and cannot be a priori neglected.

It is important to estimate the size of the various contributions to the non Wandzura-Wilczek part of $g_2$. We define the shorthand notation
\begin{align}
g_2^{\rm tw3} & = \frac{1}{2}\,\sum_a e_a^2
    \biggl(
    \widetilde g_T^{a \star}(x) 
    + \int_x^1\frac{dy}{y} \widehat{g}_T^q(y) 
    \biggr) 
&
g_2^{\text{quark}} &= \frac{1}{2}\,\sum_a e_a^2 
 \frac{\mq}{\Lambda} (h_1^q/x)^\star(x),
&
g_2^{\text{jet}} &= \frac{1}{2}\,\sum_a e_a^2 
\frac{\mj-\mq}{\Lambda} \frac{h_1^q(x)}{x}. 
\end{align} 
These are compared in Figure~\ref{f:g2contrib} to the $g_2-g_2^{WW}$ function
obtained in the very recent JAM15 fit of polarized DIS asymmetries
\cite{Sato:2016tuz}, that includes a large amount of precise data at large $x$
from Jefferson Lab, and simultaneously fits the higher-twist components of in
$g_1$ and $g_2$ to the data. For the ``pure twist-3'' contribution,
$g_2^{\rm tw3}$, {\it i.e.}, the contribution from quark-gluon-quark matrix
elements, we show a model calculation by Braun et al.~\cite{Braun:2011aw}; for
(modified) bag model calculations, see \cite{Jaffe:1990qh,Stratmann:1993aw}. To estimate the contributions
from quark ($g_2^q$) and jet mass ($g_2^{\rm jet}$) effects, that depend on chiral
odd quark-quark matrix elements, we use the recent Pavia15 fit of the
transversity distribution from Ref.~\cite{Radici:2015mwa}, which is comparable
also to other 
extractions~\cite{Anselmino:2013vqa,Kang:2015msa}. Furthermore, we choose the
values of the mass parameters to be $\mq=5$ MeV and $\mj = 100$ MeV.

As one can see, in the proton case the pure twist-3 contribution is quite
smaller in magnitude, and opposite in sign, compared to the JAM15 fit. As
expected, the quark-mass contribution is essentially negligible and cannot
reconcile these two. Even though the uncertainties in the $h_1$ extraction, and
even more in the estimate of $M_q$, are large, it is quite clear that the gap
between the pure twist-3 $g_2^{\rm tw3}$ function and the JAM15 fit can be
explained by the new jet-mass contribution we discussed in this paper.

In the neutron case, the jet contribution is very negative at intermediate to
large values of $x$. If one trusts the order of magnituse of the $g_2^{\rm tw3}$
calculation by Braun et al., one would conclude that the jet contribution
cannot be that large. However, the latter is dominated by the $d$ quark's
transversity, whose fit suffers from large systematic uncertainties and
saturates the negative Soffer bound.Recent deata in $p+p$ collisions indicate,
however, that $h_1^d$ is less negative than in the Pavia15 fits, in agreement
with the JAM15 fit of the non Wandzura-Wilczek contribution to
$g_2$. Correspondingly the jet contribution to the proton at $x \approx 0.1$
would become less positive, inproving as well the agreement with the JAM15
fit.

\begin{itemize}
\item smaller-$x$: constrains the small-$x$ behavior of transversity.
\end{itemize}
Some ideas to check:
\begin{itemize}
\item JAM13 neutron has almost 0 twist 3, with small positive contribution at
  small x. Can we use this to highlight the $g_2^{\rm jet}$ contribution? Or maybe
  as a cnstrain: if Braun is right then $g_2^{\rm jet}(n) \approx
  -g_2^{\rm tw3}(n)$. (still, does not help much for the proton). 
\item In Jam 13, Fig. 6, it looks like in general the violation of the BC sum
  rule is a small-x effect --> constraints on h(x) fits? This cannot really be
  checked vs. JAM15, that assumes $g_2^(0)$ respects BC.  
\end{itemize}

@@

The zero-th moment provides an interesting relationship between the longitudinal and tensor spin structure of the target,
\begin{align}
  \label{eq:BC}
  \int dx\, g_2(x) = \sum_q e_q^2 \frac{\mj-\mq}{\Lambda} \int dx\, \frac{1}{x} h_1^q(x) \ ,
\end{align}
where we used the fact that $\hat g_T^q[0]=0$ identically due to the symmetry properties of the quark-gluon-quark correlators. The sum rule \eqref{eq:BC}, in which the pure twist-3 part does not take part, is a consequence of the Lorentz invariance of QCD interactions, that entails $\int_0^1 dx g_1^a(x) = \int_0^1 g_T^q(x)$. It explicitly displays a contribution from spin-flip processes, that are included in the original derivation by Burkhardt and Cottingham \cite{Burkhardt:1970ti} but do not show up in treatments that only consider free field quark propagators for the struck quark \cite{Jaffe:1996zw}. Since $h_1$ is driven to 0 by QCD evolution as $Q^2 \to \infty$, the BC sum rule $\int_0^1 dx\, g_2(x) =0$ is satisfied at least asymptotically. At finite scales, assuming the Burkhardt-Cottingham sum rule, that is derived at the cross section level (under suitable assumptions on the high-energy behavior of te hspin-flip amplitude) rather than from an analysis of collinear correlators, as done here, to be strictly valid, we obtain a strong constraint on the transversity function,
\begin{align}
   \int dx\, \frac{1}{x} h_1^q(x) = 0 \ .
\label{eq:ABsumrule}
\end{align}
Interestingly, one can show that this constraint, if valid at any given scale $Q_0$ is conserved through QCD evolution. Even if the BC sum rule is broken either by a $J = 0$ fixed pole with non-polynomial residue \cite{Jaffe:1996zw} or explicitly as, \eg, in the quark target model \cite{Kundu:2001pk} -- i.e., if $g_2$ integrates to a finite but nonzero number -- we obtain that $h_1(x)/x$ must be integrable. This entails a bound on the small $x$ behavior of the transversity,
\begin{align}
  h_1^q(x) \propto x^\epsilon \ \ \ \epsilon>0 \ .
\label{eq:ABbound}
\end{align}
This bound will be very useful, \eg, in transversity fits, where the data at
small $x$ is, as yet, very limited, and in general for proper extrapolations
when calculating moments.

@@

\begin{figure}[tbh]
  \centering
  \parbox[b]{8cm}{\centering
  \includegraphics
      [width=7cm,clip=true]
      {epl_emn_h_X}
  }
  \parbox[b]{8cm}{\vskip-0.5cm\centering
  \includegraphics
      [width=7cm,clip=true]
      {epl_emn_h_h_X}
  }
  \caption{\small
	Single hadron {\it (left)} and double hadron {\it (right)}
	production in $e^+ e^-$ collisions at LO with jet and
	fragmentation correlators.
  }
  \label{fig:epl_emn_h}
\end{figure}
